---
title: "microbiome_analysis"
author: "Vienna"
date: "7/9/2021"
output: html_document
---

```{r setup, include=FALSE}
library(RSQLite)
library(DECIPHER)
library(DESeq2)
library(gridExtra)
library(ggnewscale) ##new_scale()
library(ggpubr) # ggarrange()
library(ggthemes)
library(grid) # rasterGrob()
library(vegan)
library(phyloseq)
library(png)
library(qgraph) # graph layout
library(ggsignif)
library(btools)
library(RColorBrewer)
library(tidyverse) # includes: ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(knitr)
plot_res = 300 # DPI of PNG graphs
set.seed(42)
```

```{r load}
Svtab_new <- read.delim("microbiome/data_files/feature_table.txt")
#replace . in colnames with - to match metadata.txt
colnames(Svtab_new)  <- gsub("\\.", "-", colnames(Svtab_new))
row.names(Svtab_new)<-Svtab_new[[1]] #make OTU ID the row names

#write.table(Svtab_new, file="microbiome/data_files/Svtab_new", sep='\t',col.names=TRUE, row.names=FALSE)
#open in excel, check and save as Svtab1.csv

#read in otu table
otu_table=read.csv("microbiome/data_files/Svtab1.csv",sep=",",row.names=1)
otu_table=as.matrix(otu_table)

#read in taxonomy
taxonomy=read.csv("microbiome/data_files/taxonomy.csv",sep=",",row.names=1)
taxonomy <- cbind(taxonomy, ASV = paste0("ASV", sprintf("%04d", 1:nrow(taxonomy))))
taxonomy=as.matrix(taxonomy)

metatable <- read.delim("microbiome/data_files/metadata.txt") 
#View(metatable)
row.names(metatable) <- metatable[[1]]
metatable<- metatable[,(-1)]
META <- sample_data(metatable)

#import as phyloseq objects
OTU=otu_table(otu_table,taxa_are_rows=TRUE)
TAX=tax_table(taxonomy)

sample_names(META) == sample_names(OTU)


#create phyloseq object
ps1<- phyloseq(OTU, TAX, META)

# change taxonomy header 
colnames(tax_table(ps1)) <- c(D0 = "Kingdom", D1 = "Phylum", D2 = "Class", 
                              D3 = "Order", D4 = "Family", D5 = "Genus", D6 = "Species", ASV = "ASV")

#check
rank_names(ps1)

#The total number of ASVs in the whole dataset is 
length(taxa_names(ps1))

#get rid of things we do not want
ps1 = subset_taxa(ps1, Kingdom =="Bacteria")
ps1 <- prune_taxa(taxa_sums(ps1) > 0, ps1)
ps1<-subset_taxa(ps1, (Order!="Chloroplast"))
ps1<-subset_taxa(ps1, (Family!="Mitochondria"))
length(taxa_names(ps1))
#reduces total taxa numbers from 1717 to 460

# mean, max and min of sample read counts
smin <- min(sample_sums(ps1))
smean <- mean(sample_sums(ps1))
smax <- max(sample_sums(ps1))
# printing the results
cat("The minimum sample read count is:",smin)
cat("The average sample read count is:",smean)
cat("The maximum sample read count is:",smax)

sample_sums(ps1)
```


### Plot rarefaction curves
```{r rarefaction_curve, fig.heigt=5, fig.width=12}
rare<- ps1
rare2 = subset_samples(rare, treatment2 != "positive_control")
smin <- min(sample_sums(rare2))
cat("The minimum sample read count is:",smin)

# rarefy
set.seed(42)
ps.rarefied = rarefy_even_depth(rare2, rngseed=1, sample.size=14240, replace=F)
library(ranacapa)
cbPalette <- c("#000000", "#E69F00", "#009E73", "#0072B2", "#D55E00", "#CC79A7","maroon")

ggrare(ps.rarefied, step = 100, se = FALSE, color="treatment2")+ scale_colour_manual(values=cbPalette)+geom_line(size=1)+ theme_bw()+theme(axis.line = element_line(colour = "black"))+ theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank()) + theme(text = element_text(size = 22))+theme(strip.text = element_text(face="bold", size=20))+ theme(axis.title.x=element_blank())+ggtitle("Rarefaction even depth")+theme(plot.title = element_text(hjust = 0.5, size=20, face="italic"))+theme(legend.text=element_text(size=17, face="italic"))
ggsave("microbiome/output/supp_rarefaction.png", height = 5, width = 12)
```

*** Supplementary figure: Rarefaction curves were used as a qualitative method to estimate the species richness as a function of sequencing depth for all samples. Rarefaction curves reached their asymptotes for all samples, suggesting that saturation in sequencing was achieved.

### Positive controls / mock community

The experiment contained two positive control samples with DNA from the ZymoBIOMICS™ Microbial
Community Standard (Zymo Research).


#### Plot taxa in Mock control samples

```{r mock, eval = do_tech_analysis}
ps_mock <- ps1
ps_mock <- subset_samples(ps_mock, treatment == "positive_control")
ps_mock <- prune_taxa(taxa_sums(ps_mock) > 0, ps_mock)
#transform to proportional data
ps_mock2 <- transform_sample_counts(ps_mock, function(OTU) {OTU / sum(OTU)})
ps_mock2 <- tax_glom(ps_mock2, taxrank = 'Genus')
low1pc_reads <- max(taxa_sums(ps_mock2)) * 1 / 100
low1pc_indices <- which(taxa_sums(ps_mock2) < low1pc_reads)
length(low1pc_indices)
taxa_names(ps_mock2)[low1pc_indices[1]] <- "other"
ps_merge <- merge_taxa(ps_mock2, low1pc_indices, "other")
tax_table(ps_merge)["other", ] <- c("other")

getPalette = colorRampPalette(brewer.pal(10, "Set3"))
speciesList = unique(tax_table(ps_merge)[,"Genus"])
speciesPalette = getPalette(length(speciesList))
names(speciesPalette) = speciesList

ps_df <- psmelt(ps_merge)

ps_df_sum <- ps_df %>%
  group_by(Sample,Genus) %>%
  summarise(Abundance = mean(Abundance)) %>%
  group_by(Sample) %>%
  mutate(Relative_abundance = Abundance / sum(Abundance)) %>%
  ungroup() %>%
  mutate(Genus = fct_reorder(Genus, Abundance, .fun = sum, .desc = TRUE)) %>%
  droplevels()
levels(ps_df_sum$Genus)

mock_plot<-ggplot(data = ps_df_sum, aes(x = Sample, y = Relative_abundance, fill = Genus)) +
  geom_bar(stat = "identity")+ scale_y_continuous(expand = c(0,0))+ scale_fill_manual(values= speciesPalette)+ylab("relative abundance")
#mock_plot<-mock_plot+labs(title="Mock samples")
mock_plot<-mock_plot+ theme(legend.text=element_text(size=9,face="italic"))
mock_plot<-mock_plot+theme(axis.text.x = element_text(size=12, face="bold"))+ theme(axis.title.x=element_blank())+ theme(legend.title=element_blank())+theme(axis.text.y = element_text(size=12, face="bold"))+ theme(axis.title.y=element_text(size=13, face="bold"))
mock_plot2<-mock_plot+theme(plot.title = element_text(face = "bold", color="grey40",size = (17)))+theme(plot.title = element_text(hjust = 0.5))
ps_df_sum
```

```{r figure_S1, fig.align = "center",eval = do_tech_analysis}
mock_comp <- "microbiome/data_files/mock_composition_ZymoResearch.png"
figure_S1 <- "microbiome/output/Supp_mock_composition_comparison.png"
png(figure_S1, 7 * plot_res, 4 * plot_res, res = plot_res)
ggarrange(mock_plot2, rasterGrob(readPNG(mock_comp)),labels = c("A", "B"))
invisible(dev.off())
knitr::include_graphics(figure_S1, dpi = plot_res)
```
All 8 Mock bacterial taxa correctly detected. 
**Figure S1: Comparison of the mock community with the theoretical proportions.**
The two mock samples sequenced in this study **(A)** show no big difference to the expected
theoretical proportions of the reference **(B, bar with label *theoretical*)**. The three other bacterial taxa belong to honey bee core microbiome (Bartonella, Gilliamella and Snodgrassella). The relative abundance of these three taxa across the two Mock samples account for 0.23% of all reads, representing neglectable cross-contamination during sequencing.



#### comparing difference between methods
We sequenced to types of samples: whole bee abdomen and the mix of three macerated guts for cage to cage transfers after each cycle
Therefore, we need to test if these samples are different due to the differences in the methods prior extracting before deciding if we include all or not.
Use PERMANOVA on bray-curtis dissimilarities using proportional transformed abundance data
```{r}
set.seed(42)
methods = ps1
methods <- transform_sample_counts(methods, function(OTU) {OTU / sum(OTU)})

Control = subset_samples(methods, treatment2 == "Control")
Control <- prune_taxa(taxa_sums(Control) > 0, Control)
C_metadata <- as(sample_data(Control), "data.frame")
adonis(distance(Control, method="bray") ~ sample_type,data = C_metadata,perm=999)

Tetra = subset_samples(methods, treatment2 == "Tetracycline")
Tetra <- prune_taxa(taxa_sums(Tetra) > 0, Tetra)
T_metadata <- as(sample_data(Tetra), "data.frame")
adonis(distance(Tetra, method="bray") ~ sample_type,data = T_metadata, perm=999)
```
No significant difference between gut pools or whole bees


##  Alpha diversity 

```{r alpha_Shannon, fig.heigt=13, fig.width=8}
alpha<- ps1
#pick cycle 1, 2 and 3 before stress
alpha1 = subset_samples(alpha, cycle=="cycle_one" | cycle=="cycle_two" | cycle == "cycle_three_before_stress")
#rarefy to even numbers
set.seed(1)  
alpha1_ra <- rarefy_even_depth(alpha1,sample.size=14240, replace=FALSE, rngseed = 1) 
alpha1_ra<-prune_taxa(taxa_sums(alpha1_ra) > 0, alpha1_ra)

sample_data(alpha1_ra)$cycle<-factor(sample_data(alpha1_ra)$cycle,levels=c("cycle_one","cycle_two","cycle_three_before_stress"), labels=c("Cycle one","Cycle two","Cycle three"))
levels(sample_data(alpha1_ra)$cycle)

sigFunc = function(x){
  if(x < 0.001){"***"} 
  else if(x < 0.01){"**"}
  else if(x < 0.05){"*"}
  else{NA}}

alpha_meas = c("Shannon")
p2 <- plot_richness(alpha1_ra,"treatment3",color="treatment3", measures=alpha_meas)
p2$layers <- p2$layers[-1]
p2<-p2 + geom_boxplot(data=p2$data, aes(x=treatment3, y=value),show_guide=FALSE, alpha=0.1)+geom_point(position = position_jitter(w = 0.2, h = 0.1),show_guide=FALSE) +
  geom_signif(comparisons=list(c("Control", "Tetracycline")),tip_length = 0.02,vjust=0.5, color=("grey20"),map_signif_level=sigFunc, y_position = c(3.7,3.7,3.9), margin_top = 0.1, textsize=6, test = wilcox.test)
p2 <- p2 + guides(fill=guide_legend(title=element_blank()))+ylim(1.7,4)
p2 <- p2 + theme(plot.title = element_text(hjust = 0.5))+scale_color_manual(values=c("#55596a", "#6ebe9f"))
p2 <- p2+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(panel.background = element_blank())
p2 <- p2+theme(axis.line = element_line(colour = "black"))
p2<- p2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank()) +ylab("Shannon alpha diversity")
p2 <- p2+ theme(panel.border = element_blank())
p2<-p2 + theme(text = element_text(size = 17))+theme(strip.text = element_text(face="bold", size=17, color="gray40"))+ theme(axis.title.x=element_blank())+theme(axis.text.x = element_blank())
Shannon2<-p2+ facet_grid( ~cycle,scales="free_x")+theme(axis.ticks.x=element_blank())+theme(legend.position="bottom")+theme(strip.background =element_rect(fill="white", color="white"))
Shannon2<-Shannon2+theme(panel.spacing = unit(1.8, "cm"))
Shannon2
ggsave("microbiome/output/alpha_Shannon2.png", height = 3.5, width = 6)
```
***The Shannon index  captures  information  about  both  the  species  richness  (number  of species) and the relative abundances of the species. Increased Shannon index means increased species diversity.

### make statistical comparisons on Shannon alpha diversity index - comparing treatments against respective controls in each cycle

```{r warning=FALSE}
alpha<- ps1
alpha1 = subset_samples(alpha, cycle=="cycle_one" | cycle=="cycle_two" | cycle == "cycle_three_before_stress")
#rarefy to even numbers
set.seed(1)  
alpha1_ra <- rarefy_even_depth(alpha1,sample.size=14240, replace=FALSE, rngseed = 1) 

results = estimate_richness(alpha1_ra, measures = 'Shannon')
d = sample_data(alpha1_ra)

# calculate wilcox-test
Control_1 = results[d[,'treatment_cycle3'] == 'Control_cycle_1',]
Tetracycline_1 = results[d[,'treatment_cycle3'] == 'Tetracycline_cycle_1',]
wilcox.test(Control_1, Tetracycline_1)
#W = 36, p-value = 0.004396

#cycle 2
Control_2 = results[d[,'treatment_cycle3'] == 'Control_cycle_2',]
Tetracycline_2 = results[d[,'treatment_cycle3'] == 'Tetracycline_cycle_2',]
wilcox.test(Control_2, Tetracycline_2)
#pvalue=7.396e-07)

#cycle 3
control_cycle_3_before_stress = results[d[,'treatment_cycle3'] == 'control_cycle_3_before_stress',]
Tetracycline_cycle_3_before_stress = results[d[,'treatment_cycle3'] == 'Tetracycline_cycle_3_before_stress',]
wilcox.test(control_cycle_3_before_stress, Tetracycline_cycle_3_before_stress)
#pvalue=4.114e-05
```
Under tetracycline the Shannon alpha diversity was significantly affected in all three cycles. 

### observed species numbers - alpha diversity
```{r alpha_Observed, fig.heigt=13, fig.width=8}
alpha<- ps1
alpha1 = subset_samples(alpha, treatment != "hive_bee" & treatment != "positive_control"& treatment != "start_microbiome"& cycle != "cycle_three_after_stress")
#rarefy to even numbers
smin <- min(sample_sums(alpha1))
cat("The minimum sample read count is:",smin)
set.seed(1)  
alpha1_ra <- rarefy_even_depth(alpha1,sample.size=12351, replace=FALSE, rngseed = 1) 

sample_data(alpha1_ra)$cycle<-factor(sample_data(alpha1_ra)$cycle,levels=c("cycle_one","cycle_two","cycle_three_before_stress"), labels=c("Cycle one","Cycle two","Cycle three"))
levels(sample_data(alpha1_ra)$cycle)

sigFunc = function(x){
  if(x < 0.001){"***"} 
  else if(x < 0.01){"**"}
  else if(x < 0.05){"*"}
  else{NA}}

alpha_meas = c("Observed")
p2 <- plot_richness(alpha1_ra,"treatment3",color="treatment3", measures=alpha_meas)
p2$layers <- p2$layers[-1]
p2<-p2 + geom_boxplot(data=p2$data, aes(x=treatment3, y=value),show_guide=FALSE, alpha=0.1)+geom_point(position = position_jitter(w = 0.2, h = 0.1),show_guide=FALSE) +
 geom_signif(comparisons=list(c("Control", "Tetracycline")),tip_length = 0.02,vjust=0.5, color=("grey20"),map_signif_level=sigFunc, y_position = c(73), margin_top = 0.1, textsize=6, test = wilcox.test)

p2 <- p2 + guides(fill=guide_legend(title=element_blank()))+ylim(17,84)
p2 <- p2 + theme(plot.title = element_text(hjust = 0.5))+scale_color_manual(values=c("#55596a", "#6ebe9f","#f3a935", "#D45E79"))
p2 <- p2+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(panel.background = element_blank())
p2 <- p2+theme(axis.line = element_line(colour = "black"))
p2<- p2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank()) +ylab("Observed alpha diversity")
p2 <- p2+ theme(panel.border = element_blank())
p2<-p2 + theme(text = element_text(size = 17))+theme(strip.text = element_text(face="bold", size=15, color="gray40"))+ theme(axis.title.x=element_blank())+theme(axis.text.x = element_blank())
p2<-p2+ facet_grid( ~cycle,scales="free_x",space="free")+theme(axis.ticks.x=element_blank())+theme(legend.position="bottom")+theme(strip.background =element_rect(fill="white", color="white"))
Observed<-p2+ geom_point(size = -1, shape = 15)+geom_boxplot(show.legend = FALSE)+guides(color = guide_legend(override.aes = list(size = 7,alpha=0.5)))+ theme(axis.title.y=element_text(size=14))
Observed
ggsave("microbiome/output/alpha_Observed.png", height = 3.5, width = 5)
```

*** Supplementary figure Observed alpha diversity
The observed plot shows the difference in detected OTUs. 

### make statistical comparisons on Observed species numbers - comparing treatments against respective controls in each cycle
```{r warning=FALSE}
alpha<- ps1
alpha1 = subset_samples(alpha, cycle=="cycle_one" | cycle=="cycle_two" | cycle == "cycle_three_before_stress")
#rarefy to even numbers
set.seed(1)  
alpha1_ra <- rarefy_even_depth(alpha1,sample.size=14240, replace=FALSE, rngseed = 1) 

results = estimate_richness(alpha1_ra, measures = 'Observed')
d = sample_data(alpha1_ra)

# calculate wilcox-test
Control_1 = results[d[,'treatment_cycle3'] == 'Control_cycle_1',]
Tetracycline_1 = results[d[,'treatment_cycle3'] == 'Tetracycline_cycle_1',]
wilcox.test(Control_1, Tetracycline_1)

#cycle 2
Control_2 = results[d[,'treatment_cycle3'] == 'Control_cycle_2',]
Tetracycline_2 = results[d[,'treatment_cycle3'] == 'Tetracycline_cycle_2',]
wilcox.test(Control_2, Tetracycline_2)

#cycle 3
control_cycle_3_before_stress = results[d[,'treatment_cycle3'] == 'control_cycle_3_before_stress',]
Tetracycline_cycle_3_before_stress = results[d[,'treatment_cycle3'] == 'Tetracycline_cycle_3_before_stress',]
wilcox.test(control_cycle_3_before_stress, Tetracycline_cycle_3_before_stress)
```
Numbers of observed species is significantly different under tetracycline to the control in all three cycles.



# Beta diversity

### ordination plots
```{r NMDS_man, fig.heigt=7.5, fig.width=10}
ord2<- ps1
ord2 = subset_samples(ord2, cycle=="cycle_one" | cycle=="cycle_two" | cycle == "cycle_three_before_stress")

ord2 <- transform_sample_counts(ord2, function(OTU) {OTU / sum(OTU)})
sample_data(ord2)$cycle<-factor(sample_data(ord2)$cycle,levels=c("cycle_one","cycle_two", "cycle_three_before_stress"), labels=c("Cycle one","Cycle two","Cycle three"))
levels(sample_data(ord2)$cycle)

ord2 <- prune_taxa(taxa_sums(ord2) > 0, ord2)
set.seed(42)
dist<-phyloseq::distance(ord2, method="bray") 
ordination = ordinate(ord2, method="NMDS", distance=dist)
cat("stress is:", ordination$stress)
p2 <- plot_ordination(ord2, ordination, color="treatment3") +geom_point(size=2.5)
p2 <- p2 + guides(fill=guide_legend(title=element_blank()))
p2 <- p2 + theme(plot.title = element_text(hjust = 0.5))+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p2 <- p2+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(panel.background = element_blank())
p2 <- p2+theme(axis.line = element_line(colour = "black"))
p2<- p2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank())
p2 <- p2+ theme(panel.border = element_blank())
p2<-p2 + theme(text = element_text(size = 16.5))+theme(strip.text = element_blank())
p2<-p2+ facet_wrap( ~cycle,scales="free_x")+theme(legend.position="bottom")+theme(strip.background =element_rect(fill="white", color="white"))
NMDS_main<-p2+stat_ellipse(aes(group = treatment3),size=1.1, alpha=0.5)+ theme(panel.spacing.x=unit(5.5, "lines"))+ theme(axis.title.y = element_text(margin = margin(t = 0, r = 0, b = 0, l = 20)))+ scale_x_continuous(limits = c(-0.5, 0.78), breaks = seq(-0.5, 0.78, by = 0.5))
NMDS_main
ggsave("microbiome/output/NMDS_man.png", height = 5, width = 10)
```

```{r PCoA, fig.heigt=7.5, fig.width=10}
ord2<- ps1
ord2 = subset_samples(ord2, cycle=="cycle_one" | cycle=="cycle_two" | cycle == "cycle_three_before_stress")

ord2 <- transform_sample_counts(ord2, function(OTU) {OTU / sum(OTU)})
sample_data(ord2)$cycle<-factor(sample_data(ord2)$cycle,levels=c("cycle_one","cycle_two", "cycle_three_before_stress"), labels=c("Cycle one","Cycle two","Cycle three"))
levels(sample_data(ord2)$cycle)

ord2 <- prune_taxa(taxa_sums(ord2) > 0, ord2)
set.seed(42)
dist<-phyloseq::distance(ord2, method="bray") 
ordination = ordinate(ord2, method="PCoA", distance=dist)
cat("stress is:", ordination$stress)
p2 <- plot_ordination(ord2, ordination, color="treatment3") +geom_point(size=2.5)
p2 <- p2 + guides(fill=guide_legend(title=element_blank()))
p2 <- p2 + theme(plot.title = element_text(hjust = 0.5))+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p2 <- p2+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(panel.background = element_blank())
p2 <- p2+theme(axis.line = element_line(colour = "black"))
p2<- p2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank())
p2 <- p2+ theme(panel.border = element_blank())
p2<-p2 + theme(text = element_text(size = 16.5))+theme(strip.text = element_blank())
p2<-p2+ facet_wrap( ~cycle,scales="free_x")+theme(legend.position="bottom")+theme(strip.background =element_rect(fill="white", color="white"))
PCoA_main<-p2+stat_ellipse(aes(group = treatment3),size=1.1, alpha=0.5)+ theme(panel.spacing.x=unit(5.5, "lines"))+ theme(axis.title.y = element_text(margin = margin(t = 0, r = 0, b = 0, l = 20)))+ scale_x_continuous(limits = c(-0.5, 0.78), breaks = seq(-0.5, 0.78, by = 0.5))
ggsave("microbiome/output/PCoA.png", height = 5, width = 10)
```



## Beta diversity stats

#### test difference between treatments and respective controls in the three cycles to statistically verify microbial community compositional difference seen in ordination plots 
```{r}
set.seed(42)
compare=ps1
compare1 <- transform_sample_counts(compare, function(OTU) {OTU / sum(OTU)})

cycle1 = subset_samples(compare1, cycle == "cycle_one")
cycle1 <- prune_taxa(taxa_sums(cycle1) > 0, cycle1)
cycle1.1 <- as(sample_data(cycle1), "data.frame")

cycle2 = subset_samples(compare1, cycle == "cycle_two")
cycle2 <- prune_taxa(taxa_sums(cycle2) > 0, cycle2)
cycle2.1 <- as(sample_data(cycle2), "data.frame")

cycle3b = subset_samples(compare1, cycle == "cycle_three_before_stress")
cycle3b <- prune_taxa(taxa_sums(cycle3b) > 0, cycle3b)
cycle3b.1 <- as(sample_data(cycle3b), "data.frame")

#cycle 1

subs <- subset_samples(cycle1, treatment2 %in% c("Control", "Tetracycline"))
metadata <- as(sample_data(subs), "data.frame")
adonis(distance(subs, method="bray") ~ treatment2,data = metadata, perm=999)

#cycle 2

subs <- subset_samples(cycle2, treatment2 %in% c("Control", "Tetracycline"))
metadata <- as(sample_data(subs), "data.frame")
adonis(distance(subs, method="bray") ~ treatment2,data = metadata, perm=999)

#cycle 3 before stress

subs <- subset_samples(cycle3b, treatment3 %in% c("Control", "Tetracycline"))
metadata <- as(sample_data(subs), "data.frame")
adonis(distance(subs, method="bray") ~ treatment3,data = metadata, perm=999)
```
Community composition was significantly affected under tetracycline in all cycles. 


#### Test general effects of treatment across cycles
```{r treatment_effects}
set.seed(42)
compare=ps1
compare <- prune_taxa(taxa_sums(compare) > 0, compare)
compare <- transform_sample_counts(compare, function(OTU) {OTU / sum(OTU)})
compare.1 <- as(sample_data(compare), "data.frame")

cycle1 = subset_samples(compare, cycle == "cycle_one")
cycle1 <- prune_taxa(taxa_sums(cycle1) > 0, cycle1)
cycle1.2 <- as(sample_data(cycle1), "data.frame")

cycle2 = subset_samples(compare, cycle == "cycle_two")
cycle2 <- prune_taxa(taxa_sums(cycle2) > 0, cycle2)
cycle2.2 <- as(sample_data(cycle2), "data.frame")

cycle3_before_stress = subset_samples(compare, cycle == "cycle_three_before_stress")
cycle3_before_stress <- prune_taxa(taxa_sums(cycle3_before_stress) > 0, cycle3_before_stress)
cycle3_before_stress.2 <- as(sample_data(cycle3_before_stress), "data.frame")

d = distance(compare, "bray")
adonis(d ~ treatment3, compare.1, perm=999)

d1 = distance(cycle1, "bray")
adonis(d1 ~ treatment3, cycle1.2, perm=999)

d2 = distance(cycle2, "bray")
adonis(d2 ~ treatment3, cycle2.2, perm=999)

d3 = distance(cycle3_before_stress, "bray")
adonis(d3 ~ treatment3, cycle3_before_stress.2, perm=999)
```

Treatment significantly explains differences in microbiome in all cycles (between 67 and 74 % of variation).




# combined figure alpha and beta for main figure

```{r, figure_1}

legend2 <- get_legend(PCoA_main)
legend2<-as_ggplot(legend2)
ggsave("microbiome/output/legend2.png", height = 1, width = 5)

Shannon<-Shannon2+theme(legend.position = "none")
NMDS<-NMDS_main#+theme(legend.position = "none")

library(egg) #helps to correctly align the plots

figure_1 <- "microbiome/output/alpha_and_beta.png"
png(figure_1, 11 * plot_res, 9 * plot_res, res = plot_res)
ggarrange(Shannon,NMDS,nrow=2) 
invisible(dev.off())
knitr::include_graphics(figure_1, dpi = plot_res)
```




### Test for cage effects
```{r cage_effects}
set.seed(42)

cycle1_co = subset_samples(cycle1, treatment3 == "Control")
cycle1_co.1 <- as(sample_data(cycle1_co), "data.frame")
d = distance(cycle1_co, "bray")
adonis(d ~ cage, cycle1_co.1, perm=999)

#cycle2

cycle2_control = subset_samples(cycle2, treatment3 == "Control")
cycle2_control.1 <- as(sample_data(cycle2_control), "data.frame")
d = distance(cycle2_control, "bray")
adonis(d ~ cage, cycle2_control.1, perm=999)

cycle2_tet = subset_samples(cycle2, treatment3 == "Tetracycline")
cycle2_tet.1 <- as(sample_data(cycle2_tet), "data.frame")
d = distance(cycle2_tet, "bray")
adonis(d ~ cage, cycle2_tet.1, perm=999)

#cycle 3 before stress

cycle3_control = subset_samples(cycle3_before_stress, treatment3 == "Control")
cycle3_control.1 <- as(sample_data(cycle3_control), "data.frame")
d = distance(cycle3_control, "bray")
adonis(d ~ cage, cycle3_control.1, perm=999)

cycle3_tet = subset_samples(cycle3_before_stress, treatment3 == "Tetracycline")
cycle3_tet.1 <- as(sample_data(cycle3_tet), "data.frame")
d = distance(cycle3_tet, "bray")
adonis(d ~ cage, cycle3_tet.1, perm=999)
```

Testing if the three cages within a treatment show significant variations in microbiome composition (cage effects) in cycle 2 control and cycle 3 control significant cage variations are seen.


### Time effect
#### Test if microbial composition of treatments differ across the three cycles

```{r}
set.seed(42)
compare = ps1
compare2 <- prune_taxa(taxa_sums(compare) > 0, compare)
compare2 <- transform_sample_counts(compare2, function(OTU) {OTU / sum(OTU)})

subs1 <- subset_samples(compare2, treatment_cycle3 == "Control_cycle_1"|treatment_cycle3=="Control_cycle_2"|treatment_cycle3=="control_cycle_3_before_stress")
metadata <- as(sample_data(subs1), "data.frame")
adonis(distance(subs1, method="bray") ~ cycle,
       data = metadata, perm=999)

subs4 <- subset_samples(compare2, treatment_cycle %in% c("Tetracycline_cycle_1", "Tetracycline_cycle_2","Tetracycline_cycle_3_before_stress"))
metadata <- as(sample_data(subs4), "data.frame")
adonis(distance(subs4, method="bray") ~ cycle,
       data = metadata, perm=999)
```

Treatments differ significantly across cycles

#### compare treatments before and after high stress

```{r}
set.seed(42)
compare2 = ps1
compare2 <- transform_sample_counts(compare2, function(OTU) {OTU / sum(OTU)})

subs1 <- subset_samples(compare2, treatment_cycle %in% c("control_cycle_3_before_stress", "control_Tetracycline_cycle_3_after_stress"))
metadata <- as(sample_data(subs1), "data.frame")
adonis(distance(subs1, method="bray") ~ treatment_cycle,data = metadata, perm=999)

subs1 <- subset_samples(compare2, treatment_cycle %in% c("Tetracycline_cycle_3_before_stress", "Tetracycline_cycle_3_after_stress"))
metadata <- as(sample_data(subs1), "data.frame")
adonis(distance(subs1, method="bray") ~ treatment_cycle,data = metadata, perm=999)
```

No significant difference for any treatment before and after high stress application -> likely the time was not enough in between the two time points and we sequenced a lot of dead bacteria


### test spread of variance of groups
```{r betadisp}
set.seed(53)
test<-ps1
test2 <- transform_sample_counts(test, function(OTU) {OTU / sum(OTU)})

cycle1 = subset_samples(test2, cycle == "cycle_one")
cycle1 <- prune_taxa(taxa_sums(cycle1) > 0, cycle1)

cycle2 = subset_samples(test2, cycle == "cycle_two")
cycle2 <- prune_taxa(taxa_sums(cycle2) > 0, cycle2)

cycle3_b = subset_samples(test2, cycle == "cycle_three_before_stress")
cycle3_b <- prune_taxa(taxa_sums(cycle3_b) > 0, cycle3_b)

#testing treatment
set.seed(53)
d_cycle1 = distance(cycle1, "bray")
df_cycle1 = as(sample_data(cycle1), "data.frame")
groups <- df_cycle1[["treatment3"]]
beta <- betadisper(d_cycle1, df_cycle1$treatment3)
permutest(beta,pairwise = TRUE, permutations = 999)
anova(betadisper(d_cycle1, groups))
mod.HSD <- TukeyHSD(beta,conf.level = 0.95)
mod.HSD
plot(betadisper(d_cycle1, groups),main="MultiVariate Permutation Cycle one",lwd=2, label = TRUE,label.cex = 0.7,col=c("#55596a", "#6ebe9f"))
boxplot(betadisper(d_cycle1, groups),main="Cycle one",border=c("#55596a", "#6ebe9f"),xlab="")

set.seed(53)
d_cycle2 = distance(cycle2, "bray")
df_cycle2 = as(sample_data(cycle2), "data.frame")
groups <- df_cycle2[["treatment3"]]
beta <- betadisper(d_cycle2, df_cycle2$treatment3)
permutest(beta,pairwise = TRUE, permutations = 99)
anova(betadisper(d_cycle2, groups))
mod.HSD <- TukeyHSD(beta)
mod.HSD
plot(betadisper(d_cycle2, groups),main="MultiVariate Permutation Cycle two",lwd=2, label = TRUE,label.cex = 0.7,col=c("#55596a", "#6ebe9f"))
boxplot(betadisper(d_cycle2, groups),main="Cycle two",border=c("#55596a", "#6ebe9f"),xlab="")

set.seed(53)
d_cycle3_b = distance(cycle3_b, "bray")
df_cycle3_b = as(sample_data(cycle3_b), "data.frame")
groups <- df_cycle3_b[["treatment3"]]
beta <- betadisper(d_cycle3_b, df_cycle3_b$treatment3)
permutest(beta,pairwise = TRUE, permutations = 99)
anova(betadisper(d_cycle3_b, groups))
mod.HSD <- TukeyHSD(beta)
mod.HSD
plot(betadisper(d_cycle3_b, groups),main="MultiVariate Permutation Cycle three",lwd=2, label = TRUE,label.cex = 0.7,col=c("#55596a", "#6ebe9f"))
boxplot(betadisper(d_cycle3_b, groups),main="Cycle three",border=c("#55596a", "#6ebe9f"),xlab="")
```

# Taxonomy plots and stats

```{r set_up_for_taxa_plots}
tax <-ps1
tax = subset_samples(tax, cycle == "cycle_one"| cycle == "cycle_two"| cycle == "cycle_three_before_stress")
ps_bdiv2 <- transform_sample_counts(tax, function(OTU) {OTU / sum(OTU)})
tax_all2<- ps_bdiv2
ps_pd2 <- tax_glom(tax_all2, taxrank = 'Genus')

low1pc_reads <- max(taxa_sums(ps_pd2)) * 1 / 100
low1pc_indices <- which(taxa_sums(ps_pd2) < low1pc_reads)
length(low1pc_indices)
taxa_names(ps_pd2)[low1pc_indices[1]] <- "other"
taxa3 <- merge_taxa(ps_pd2, low1pc_indices, "other")
tax_table(taxa3)["other", ] <- c("other")

#build a custom color palette for phyloseq object
getPalette = colorRampPalette(brewer.pal(8, "Set3"))
speciesList = unique(tax_table(taxa3)[,"Genus"])
speciesPalette = getPalette(length(speciesList))
names(speciesPalette) = speciesList
```


```{r taxa_cycle1}

ps2 <-taxa3
cycle1 = subset_samples(ps2, cycle=="cycle_one")
cycle1 <- prune_taxa(taxa_sums(cycle1) > 0, cycle1)

ps_df_c1 <- psmelt(cycle1)
ps_df_sum_c1 <- ps_df_c1 %>%
  group_by(treatment3,Genus,cycle,sample) %>%
  summarise(Abundance = mean(Abundance)) %>%
  group_by(cycle,treatment3,sample) %>%
  mutate(relative_abundance = Abundance / sum(Abundance)) %>%
  ungroup() %>%
  mutate(Genus = fct_reorder(Genus, Abundance, .fun = sum, .desc = TRUE))
levels(ps_df_sum_c1$Genus)

change <- c("Cycle one")
ps_df_sum_c1 <- arrange(transform(ps_df_sum_c1, cycle=factor(cycle,labels=change)),cycle)

ps_df_sum_c1$sample <- factor(ps_df_sum_c1$sample, levels = unique(ps_df_sum_c1$sample[order(ps_df_sum_c1$treatment3)]))

p<-ggplot(data = ps_df_sum_c1, aes(x = sample, y = relative_abundance, fill = Genus)) +
  geom_bar(stat = "identity")+ scale_fill_manual(values= speciesPalette)+ scale_y_continuous(expand = c(0,0))+ylab("relative abundance")
p<-p+ theme(legend.text=element_text(size=18, face = "italic"))+theme(legend.key = element_rect(color = NA, fill = NA),
                                                     legend.key.size = unit(0.9, "cm"))+theme(legend.title = element_blank())+theme(axis.ticks.x = element_blank())
p<-p + facet_grid(cycle~treatment3, scales="free_x",space="free_x")
p<-p + theme(strip.text = element_text(size=18,face="bold",color="grey35"))
cycle1_plot<-p+theme(axis.title.y = element_text(size=17, face="bold"))+theme(axis.text.y = element_text(size=15, face="bold"))+theme(axis.text.x = element_text(size=11, face="bold", angle = 90))+theme(axis.title.x = element_blank())+theme(strip.background =element_rect(fill="white",color="white"))+theme(axis.text.x = element_blank())+theme(legend.position="bottom")
ggsave("microbiome/output/cycle1_plot.png", height = 6, width = 10)
```

```{r taxa_cycle2}

ps2 <-taxa3
cycle2 = subset_samples(ps2, cycle=="cycle_two")
cycle2 <- prune_taxa(taxa_sums(cycle2) > 0, cycle2)

ps_df_c2 <- psmelt(cycle2)
ps_df_sum_c2 <- ps_df_c2 %>%
  group_by(treatment3,Genus,cycle,sample) %>%
  summarise(Abundance = mean(Abundance)) %>%
  group_by(cycle,treatment3,sample) %>%
  mutate(relative_abundance = Abundance / sum(Abundance)) %>%
  ungroup() %>%
  mutate(Genus = fct_reorder(Genus, Abundance, .fun = sum, .desc = TRUE))
levels(ps_df_sum_c2$Genus)

change <- c("Cycle two")
ps_df_sum_c2 <- arrange(transform(ps_df_sum_c2, cycle=factor(cycle,labels=change)),cycle)

ps_df_sum_c2$sample <- factor(ps_df_sum_c2$sample, levels = unique(ps_df_sum_c2$sample[order(ps_df_sum_c2$treatment3)]))

p<-ggplot(data = ps_df_sum_c2, aes(x = sample, y = relative_abundance, fill = Genus)) +
  geom_bar(stat = "identity")+ scale_fill_manual(values= speciesPalette)+ scale_y_continuous(expand = c(0,0))+ylab("relative abundance")
p<-p+ theme(legend.text=element_text(size=18, face = "italic"))+theme(legend.key = element_rect(color = NA, fill = NA),
                                                     legend.key.size = unit(0.9, "cm"))+theme(legend.title = element_blank())
#p<-p + facet_wrap(cycle~treatment3, scales="free_x",nrow = 1)
p<-p + facet_grid(cycle~treatment3, scales="free_x",space="free_x")
p<-p + theme(strip.text = element_text(size=18,face="bold",color="grey35"))
cycle2_plot<-p+theme(axis.title.y = element_text(size=18, face="bold"))+theme(axis.text.y = element_text(size=15, face="bold"))+theme(axis.text.x = element_text(size=11, face="bold", angle = 90))+theme(axis.title.x = element_blank())+theme(axis.text.x = element_text(vjust = 0.5,hjust=1))+theme(strip.background =element_rect(fill="white",color="white"))+theme(legend.position="bottom")+ theme(axis.text.x = element_text(hjust = -0.2))+theme(axis.text.x = element_blank())+theme(axis.ticks.x = element_blank())
ggsave("microbiome/output/cycle2_plot.png", height = 6, width = 10)
```


```{r taxa_cycle3}

ps2 <-taxa3
cycle3 = subset_samples(ps2, cycle=="cycle_three_before_stress")
cycle3 <- prune_taxa(taxa_sums(cycle3) > 0, cycle3)

ps_df_c3 <- psmelt(cycle3)
ps_df_sum_c3 <- ps_df_c3 %>%
  group_by(treatment3,Genus,cycle,sample) %>%
  summarise(Abundance = mean(Abundance)) %>%
  group_by(cycle,treatment3,sample) %>%
  mutate(relative_abundance = Abundance / sum(Abundance)) %>%
  ungroup() %>%
  mutate(Genus = fct_reorder(Genus, Abundance, .fun = sum, .desc = TRUE))
levels(ps_df_sum_c3$Genus)

change <- c("Cycle three")
ps_df_sum_c3 <- arrange(transform(ps_df_sum_c3, cycle=factor(cycle,labels=change)),cycle)

ps_df_sum_c3$sample <- factor(ps_df_sum_c3$sample, levels = unique(ps_df_sum_c3$sample[order(ps_df_sum_c3$treatment3)]))

p<-ggplot(data = ps_df_sum_c3, aes(x = sample, y = relative_abundance, fill = Genus)) +
  geom_bar(stat = "identity")+ scale_fill_manual(values= speciesPalette)+ scale_y_continuous(expand = c(0,0))+ylab("relative abundance")
p<-p+ theme(legend.text=element_text(size=18, face = "italic"))+theme(legend.key = element_rect(color = NA, fill = NA),
                                                     legend.key.size = unit(0.7, "cm"))+theme(legend.title = element_blank())
#p<-p + facet_wrap(cycle~treatment3, scales="free_x",nrow = 1)
p<-p + facet_grid(cycle~treatment3, scales="free_x",space="free_x")
p<-p + theme(strip.text = element_text(size=18,face="bold",color="grey35"))
cycle3_plot<-p+theme(axis.title.y = element_text(size=17, face="bold"))+theme(axis.text.y = element_text(size=15, face="bold"))+theme(axis.text.x = element_text(size=11, face="bold", angle = 90))+theme(axis.title.x = element_blank())+theme(axis.text.x = element_text(vjust = 0.5,hjust=1))+theme(strip.background =element_rect(fill="white",color="white"))+theme(legend.position="bottom")+ theme(axis.text.x = element_text(hjust = -0.2))+theme(axis.text.x = element_blank())+theme(axis.ticks.x = element_blank())
ggsave("microbiome/output/cycle3_plot.png", height = 6, width = 10)
```

```{r taxa_plot_all_combine}

legend <- get_legend(cycle3_plot)
legend<-as_ggplot(legend)
ggsave("microbiome/output/legend.png", height = 1, width = 11)

cycle1<-cycle1_plot+theme(legend.position = "none")+theme(axis.title.y = element_blank())
cycle2<-cycle2_plot+theme(legend.position = "none")
cycle3<-cycle3_plot+theme(axis.title.y = element_blank())+theme(legend.position = "none")

library(egg) #helps to correctly align the plots

figure_1 <- "microbiome/output/taxa_plot_all.png"
png(figure_1, 8.1 * plot_res, 9.5 * plot_res, res = plot_res)
ggarrange(cycle1,cycle2,cycle3,nrow=3) 
invisible(dev.off())
knitr::include_graphics(figure_1, dpi = plot_res)

```

### stats on abundances

### compare if taxa abundances are significantly different between controls and treatments in cycle 1, 2 and cycle 3, use Wilcoxon tests with following fdr correction. In addition test if control samples differ in the abundances of the core taxa between cycle 1 and 3 (time or lab-adaptation effect)

```{r, lactobacillus_differ}
set.seed(42)
ps2 <-ps1
ps2 <- rarefy_even_depth(ps2,sample.size=14240, replace=FALSE, rngseed = 1) 
ps2 = subset_samples(ps2, cycle == "cycle_one"| cycle == "cycle_two"| cycle == "cycle_three_before_stress")
ps3 <- prune_taxa(taxa_sums(ps2) > 0, ps2)
abundance <- tax_glom(ps3, taxrank = 'Genus')
Lac = subset_taxa(abundance, Genus=="Lactobacillus")

#cycle 1
Tet <- subset_samples(Lac, treatment_cycle =="Control_cycle_1" | treatment_cycle =="Tetracycline_cycle_1")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

#cycle 2
Tet <- subset_samples(Lac, treatment_cycle =="Control_cycle_2" | treatment_cycle =="Tetracycline_cycle_2")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

#cycle 3
Tet <- subset_samples(Lac, treatment_cycle=="control_cycle_3_before_stress"|treatment_cycle=="Tetracycline_cycle_3_before_stress")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment3, data=melt)
```

```{r, Bartonella_differ}
Bart = subset_taxa(abundance, Genus=="Bartonella")

Tet <- subset_samples(Bart, treatment_cycle =="Control_cycle_1" | treatment_cycle =="Tetracycline_cycle_1")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Bart, treatment_cycle =="Control_cycle_2" | treatment_cycle =="Tetracycline_cycle_2")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Bart, treatment_cycle=="control_cycle_3_before_stress"|treatment_cycle=="Tetracycline_cycle_3_before_stress")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment3, data=melt)
```
#### Gilliamella
```{r,Gilliamella_abund}
Gil = subset_taxa(abundance, Genus=="Gilliamella")

Tet <- subset_samples(Gil, treatment_cycle =="Control_cycle_1" | treatment_cycle =="Tetracycline_cycle_1")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Gil, treatment_cycle =="Control_cycle_2" | treatment_cycle =="Tetracycline_cycle_2")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Gil, treatment_cycle=="control_cycle_3_before_stress"|treatment_cycle=="Tetracycline_cycle_3_before_stress")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment3, data=melt)
```
#### Snodgrassella
```{r,Snodgrassella_abund}
Snod = subset_taxa(abundance, Genus=="Snodgrassella")

Tet <- subset_samples(Snod, treatment_cycle =="Control_cycle_1" | treatment_cycle =="Tetracycline_cycle_1")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Snod, treatment_cycle =="Control_cycle_2" | treatment_cycle =="Tetracycline_cycle_2")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Snod, treatment_cycle=="control_cycle_3_before_stress"|treatment_cycle=="Tetracycline_cycle_3_before_stress")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment3, data=melt)
```


#### Commensalibacter
```{r,Commensalibacter_abund}
Com = subset_taxa(abundance, Genus=="Commensalibacter")

Tet <- subset_samples(Com, treatment_cycle =="Control_cycle_1" | treatment_cycle =="Tetracycline_cycle_1")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Com, treatment_cycle =="Control_cycle_2" | treatment_cycle =="Tetracycline_cycle_2")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Com, treatment_cycle=="control_cycle_3_before_stress"|treatment_cycle=="Tetracycline_cycle_3_before_stress")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment3, data=melt)
```
#### Bifidobacterium
```{r,Bifidobacterium_abund}
Bif = subset_taxa(abundance, Genus=="Bifidobacterium")

Tet <- subset_samples(Bif, treatment_cycle =="Control_cycle_1" | treatment_cycle =="Tetracycline_cycle_1")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Bif, treatment_cycle =="Control_cycle_2" | treatment_cycle =="Tetracycline_cycle_2")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment2, data=melt)

Tet <- subset_samples(Bif, treatment_cycle=="control_cycle_3_before_stress"|treatment_cycle=="Tetracycline_cycle_3_before_stress")
Tet <- prune_taxa(taxa_sums(Tet) > 0, Tet)
melt<-psmelt(Tet)
wilcox.test(Abundance~treatment3, data=melt)

```


## Core taxa abundance

```{r total abundance}
ps2 <-ps1
ps2 <- rarefy_even_depth(ps2,sample.size=14240, replace=FALSE, rngseed = 1) 
ps2 = subset_samples(ps2, cycle == "cycle_one"| cycle == "cycle_two"| cycle == "cycle_three_before_stress")
ps3 <- prune_taxa(taxa_sums(ps2) > 0, ps2)
ps3 <- tax_glom(ps3, taxrank = 'Genus')
psOrd3 = subset_taxa(ps3, Genus=="Lactobacillus" | Genus=="Bartonella" | Genus=="Gilliamella" | Genus=="Snodgrassella" | Genus=="Frischella" | Genus=="Bifidobacterium" | Genus=="Commensalibacter")

#Melt and plot
melt<-psmelt(psOrd3)
levels2=c("cycle_one","cycle_two","cycle_three_before_stress")
change <- c("Cycle one","Cycle two","Cycle three")
melt2 <- arrange(transform(melt, cycle=factor(cycle,levels=levels2,labels=change)),cycle)
neworder2 <- c("Bartonella","Lactobacillus","Gilliamella","Snodgrassella","Bifidobacterium","Commensalibacter","Frischella")
melt4 <- arrange(transform(melt2, Genus=factor(Genus,levels=neworder2)),Genus)

a_mean <- melt4 %>% 
  group_by(treatment3,Genus,cycle) %>% 
  summarize(mean_val = mean(Abundance))
a_mean2<-subset(a_mean, treatment3 =="Control")
print(a_mean2)

sigFunc = function(x){
  if(x < 0.001){"***"} 
  else if(x < 0.01){"**"}
  else if(x < 0.05){"*"}
  else{NA}}

p<-ggplot(data = melt4, aes(x = treatment3, y = Abundance)) +
  geom_boxplot(aes(fill=Genus),alpha=0.5,lwd=0.7, position = position_dodge(width = 0.3), width=0.45,outlier.shape = NA) + geom_jitter(aes(color = Genus), height = 0, width = .1)+geom_hline(data= a_mean2, aes(yintercept=mean_val), width=5,size=1,linetype="dashed")+
  labs(x = "", y = "Abundance\n")+
  facet_grid(Genus~cycle, scales = "free")+theme_bw()+
  geom_signif(comparisons=list(c("Control", "Tetracycline")),tip_length = 0.03,vjust=1.3, color="grey20", size=1.1,map_signif_level=sigFunc, textsize=9, test = wilcox.test,step_increase = 0.2)
p<-p+ theme(legend.position="right")+ylab("Total abundance")
p<-p+ theme(legend.text=element_text(size=29, face = "italic"))+theme(legend.key = element_rect(color = NA, fill = NA),legend.key.size = unit(0.9, "cm"))+theme(legend.title = element_blank())
abu<-p +theme(strip.text.x = element_text(size=25,face="bold",color = "grey 35"))+theme(strip.background =element_rect(fill="white", color="white"))
cycle<-abu+theme(axis.title.y = element_text(size=25, face="bold"))+theme(axis.text.y = element_text(size=20, face="bold"))+theme(axis.text.x = element_text(size=25, angle=90,face="bold"))+theme(axis.title.x = element_blank())+ theme(strip.text.y = element_blank())
ggsave("microbiome/output/core_total_abundance.png", height = 20, width = 14)
``` 



### ASV


### have a look on most abundant ASVs in the core taxa

```{r Gilliamella_ASV}
tax<-ps1
tax = rarefy_even_depth(tax, rngseed=1, sample.size=14240, replace=F)
ps2 <- subset_samples(tax, cycle =="cycle_one"|cycle =="cycle_two"|cycle =="cycle_three_before_stress")
ps2 <- subset_samples(ps2, treatment3 =="Control"|treatment3 =="Tetracycline")
ps2 <- subset_taxa(ps2, Genus =="Gilliamella")

ps2.3 <- prune_taxa(taxa_sums(ps2) > 1000, ps2)
ps2.4 <- subset_taxa(ps2.3, ASV !="ASV0068")

ps2m <- psmelt(ps2.4)

neworder <- c("cycle_one","cycle_two","cycle_three_before_stress")
change <- c("Cycle one","Cycle two","Cycle three")
ps_df_sum2 <- arrange(transform(ps2m, cycle=factor(cycle,levels=neworder,labels=change)),cycle)

p<-ggplot(data = ps_df_sum2, aes(x = treatment3, y = Abundance, colour = treatment3)) +
  geom_point(position = position_jitter())+geom_boxplot(show.legend=FALSE, alpha = 0,colour="grey53",size=0.7)+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p<-p + facet_wrap(~ASV, scales="free_y",nrow=2)
p<-p + theme(strip.text = element_text(size=13,face="bold",color="grey40"))
taxa2<-p+theme(axis.title.y = element_text(size=13, face="bold"))+theme(axis.text.y = element_text(size=12, face="bold"))+theme(axis.text.x = element_blank())+theme(axis.title.x = element_blank())+theme(strip.background =element_rect(fill="white", color="white"))
taxa2<- taxa2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank(),legend.text = element_text(size=14, face="italic"))
taxa2+ggtitle("Gilliamella apicola")+theme(plot.title = element_text(hjust = 0.5, size=14, face="bold.italic",colour = "grey38"))
ggsave("microbiome/output/ASV_Gilliamella.png", height = 4.5, width = 8)
```


```{r Bifidobacterium_ASV}
tax<-ps1
tax = rarefy_even_depth(tax, rngseed=1, sample.size=14240, replace=F)
ps2 <- subset_samples(tax, cycle =="cycle_one"|cycle =="cycle_two"|cycle =="cycle_three_before_stress")
ps2 <- subset_samples(ps2, treatment3 =="Control"|treatment3 =="Tetracycline")
ps2 <- subset_taxa(ps2, Genus =="Bifidobacterium")

ps2.3 <- prune_taxa(taxa_sums(ps2) > 1000, ps2)
ps2m <- psmelt(ps2.3)

neworder <- c("cycle_one","cycle_two","cycle_three_before_stress")
change <- c("Cycle one","Cycle two","Cycle three")
ps_df_sum2 <- arrange(transform(ps2m, cycle=factor(cycle,levels=neworder,labels=change)),cycle)

p<-ggplot(data = ps_df_sum2, aes(x = treatment3, y = Abundance, colour = treatment3)) +
  geom_point(position = position_jitter())+geom_boxplot(show.legend=FALSE, alpha = 0,colour="grey53",size=0.7)+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p<-p + facet_wrap(~ASV, scales="free_y",nrow=2)
p<-p + theme(strip.text = element_text(size=13,face="bold",color="grey40"))
taxa2<-p+theme(axis.title.y = element_text(size=13, face="bold"))+theme(axis.text.y = element_text(size=12, face="bold"))+theme(axis.text.x = element_blank())+theme(axis.title.x = element_blank())+theme(strip.background =element_rect(fill="white", color="white"))
taxa2<- taxa2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank(),legend.text = element_text(size=14, face="italic"))
taxa2+ggtitle("Bifidobacterium sp.")+theme(plot.title = element_text(hjust = 0.5, size=14, face="bold.italic",colour = "grey38"))
ggsave("microbiome/output/ASV_Bifidobacterium.png", height = 4.5, width = 7)
```

### species Lactobacillus
```{r Lacto_species}
tax<-ps1
tax = rarefy_even_depth(tax, rngseed=1, sample.size=14240, replace=F)
ps2 <- subset_samples(tax, cycle =="cycle_one"|cycle =="cycle_two"|cycle =="cycle_three_before_stress")
ps2 <- subset_samples(ps2, treatment3 =="Control"|treatment3 =="Tetracycline")
ps2 <- subset_taxa(ps2, Genus =="Lactobacillus")
ps2 <- prune_taxa(taxa_sums(ps2) > 1000, ps2)
ps2 <- tax_glom(ps2, taxrank = 'Species')
ps2m <- psmelt(ps2)

neworder <- c("cycle_one","cycle_two","cycle_three_before_stress")
change <- c("Cycle one","Cycle two","Cycle three")
ps_df_sum2 <- arrange(transform(ps2m, cycle=factor(cycle,levels=neworder,labels=change)),cycle)

neworder3 <- c("Lactobacillus apis","Lactobacillus kullabergensis","Lactobacillus helsingborgensis","Lactobacillus kunkeei","Lactobacillus sp. G3_4_3CO2","Lactobacillus melliventris","not_available")
change3 <- c("L. apis","L. kullabergensis","L. helsingborgensis","L. kunkeei","L. kimbladii","L. melliventis","not_available")
ps_df_sum4 <- arrange(transform(ps_df_sum2, Species=factor(Species,levels=neworder3,labels=change3)),Species)

p<-ggplot(data = ps_df_sum4, aes(x = treatment3, y = Abundance, colour = treatment3)) +
  geom_point(position = position_jitter())+geom_boxplot(show.legend=FALSE, alpha = 0,colour="grey53",size=0.7)+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p<-p + facet_wrap(~Species, scales="free_y",nrow=3)
p<-p + theme(strip.text = element_text(size=13,face="bold",color="grey40"))
taxa2<-p+theme(axis.title.y = element_text(size=13, face="bold"))+theme(axis.text.y = element_text(size=12, face="bold"))+theme(axis.text.x = element_blank())+theme(axis.title.x = element_blank())+theme(strip.background =element_rect(fill="white", color="white"))
taxa2<- taxa2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank(),legend.text = element_text(size=14, face="italic"))
taxa2+ggtitle("Lactobacillus spp.")+theme(plot.title = element_text(hjust = 0.5, size=14, face="bold.italic",colour = "grey38"))
ggsave("microbiome/output/Lactobacillus spp.png", height = 7, width = 11)
```


### ASV Lactobacillus
```{r ASV_L.apis}
tax<-ps1
tax = rarefy_even_depth(tax, rngseed=1, sample.size=14240, replace=F)
ps2 <- subset_samples(tax, cycle =="cycle_one"|cycle =="cycle_two"|cycle =="cycle_three_before_stress")
ps2 <- subset_samples(ps2, treatment3 =="Control"|treatment3 =="Tetracycline")
ps2 <- subset_taxa(ps2, Genus =="Lactobacillus")
ps2 <- subset_taxa(ps2, Species =="Lactobacillus apis")
ps2.3 <- prune_taxa(taxa_sums(ps2) > 1000, ps2)
ps2.3 <- prune_taxa(taxa_sums(ps2.3) > 0, ps2.3)
ps2m <- psmelt(ps2.3)
neworder <- c("cycle_one","cycle_two","cycle_three_before_stress")
change <- c("Cycle one","Cycle two","Cycle three")
ps_df_sum2 <- arrange(transform(ps2m, cycle=factor(cycle,levels=neworder,labels=change)),cycle)

p<-ggplot(data = ps_df_sum2, aes(x = treatment3, y = Abundance, colour = treatment3)) +
  geom_point(position = position_jitter())+geom_boxplot(show.legend=FALSE, alpha = 0,colour="grey53",size=0.7)+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p<-p + facet_wrap(~ASV, scales="free_y",nrow=1)
p<-p + theme(strip.text = element_text(size=13,face="bold",color="grey40"))
taxa2<-p+theme(axis.title.y = element_text(size=13, face="bold"))+theme(axis.text.y = element_text(size=12, face="bold"))+theme(axis.text.x = element_blank())+theme(axis.title.x = element_blank())+theme(strip.background =element_rect(fill="white", color="white"))
taxa2<- taxa2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank(),legend.text = element_text(size=14, face="italic"))
taxa2+ggtitle("Lactobacillus apis")+theme(plot.title = element_text(hjust = 0.5, size=14, face="bold.italic",colour = "grey38"))
ggsave("microbiome/output/ASV_Lactobacillus apis.png", height = 3, width = 8)
```
# Lactobacillus kullabergensis
```{r ASV_L.kullabergensis}
tax<-ps1
tax = rarefy_even_depth(tax, rngseed=1, sample.size=14240, replace=F)
ps2 <- subset_samples(tax, cycle =="cycle_one"|cycle =="cycle_two"|cycle =="cycle_three_before_stress")
ps2 <- subset_samples(ps2, treatment3 =="Control"|treatment3 =="Tetracycline")
ps2 <- subset_taxa(ps2, Genus =="Lactobacillus")
ps2 <- subset_taxa(ps2, Species =="Lactobacillus kullabergensis")
ps2.3 <- prune_taxa(taxa_sums(ps2) > 1000, ps2)
ps2.3 <- prune_taxa(taxa_sums(ps2.3) > 0, ps2.3)
ps2m <- psmelt(ps2.3)
neworder <- c("cycle_one","cycle_two","cycle_three_before_stress")
change <- c("Cycle one","Cycle two","Cycle three")
ps_df_sum2 <- arrange(transform(ps2m, cycle=factor(cycle,levels=neworder,labels=change)),cycle)

p<-ggplot(data = ps_df_sum2, aes(x = treatment3, y = Abundance, colour = treatment3)) +
  geom_point(position = position_jitter())+geom_boxplot(show.legend=FALSE, alpha = 0,colour="grey53",size=0.7)+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p<-p + facet_wrap(~ASV, scales="free_y",nrow=1)
p<-p + theme(strip.text = element_text(size=13,face="bold",color="grey40"))
taxa2<-p+theme(axis.title.y = element_text(size=13, face="bold"))+theme(axis.text.y = element_text(size=12, face="bold"))+theme(axis.text.x = element_blank())+theme(axis.title.x = element_blank())+theme(strip.background =element_rect(fill="white", color="white"))
taxa2<- taxa2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank(),legend.text = element_text(size=14, face="italic"))
taxa2+ggtitle("Lactobacillus kullabergensis")+theme(plot.title = element_text(hjust = 0.5, size=14, face="bold.italic",colour = "grey38"))
ggsave("microbiome/output/ASV_Lactobacillus kullabergensis.png", height = 3, width = 8)
```

# Lactobacillus NA
```{r ASV_L.NA}
tax<-ps1
tax = rarefy_even_depth(tax, rngseed=1, sample.size=14240, replace=F)
ps2 <- subset_samples(tax, cycle =="cycle_one"|cycle =="cycle_two"|cycle =="cycle_three_before_stress")
ps2 <- subset_samples(ps2, treatment3 =="Control"|treatment3 =="Tetracycline")
ps2 <- subset_taxa(ps2, Genus =="Lactobacillus")
ps2 <- subset_taxa(ps2, Species =="not_available")
#ps2 <- subset_taxa(ps2, ASV !="ASV0069")
ps2.3 <- prune_taxa(taxa_sums(ps2) > 1000, ps2)
ps2m <- psmelt(ps2.3)
neworder <- c("cycle_one","cycle_two","cycle_three_before_stress")
change <- c("Cycle one","Cycle two","Cycle three")
ps_df_sum2 <- arrange(transform(ps2m, cycle=factor(cycle,levels=neworder,labels=change)),cycle)

#neworder3 <- c("ASV0004","ASV0005","ASV0009", "ASV0027","ASV0031","ASV0045","ASV0052","ASV0067","ASV0030","ASV0033","ASV0035","ASV0036","ASV0043","ASV0044","ASV0063","ASV0019","ASV0037")
#ps_df_sum4 <- arrange(transform(ps_df_sum2, ASV=factor(ASV,levels=neworder3)),ASV)

p<-ggplot(data = ps_df_sum2, aes(x = treatment3, y = Abundance, colour = treatment3)) +
  geom_point(position = position_jitter())+geom_boxplot(show.legend=FALSE, alpha = 0,colour="grey53",size=0.7)+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p<-p + facet_wrap(~ASV, scales="free_y",nrow=4)
p<-p + theme(strip.text = element_text(size=13,face="bold",color="grey40"))
taxa2<-p+theme(axis.title.y = element_text(size=13, face="bold"))+theme(axis.text.y = element_text(size=12, face="bold"))+theme(axis.text.x = element_blank())+theme(axis.title.x = element_blank())+theme(strip.background =element_rect(fill="white", color="white"))
taxa2<- taxa2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank(),legend.text = element_text(size=14, face="italic"))
taxa2+ggtitle("Lactobacillus sp. NA")+theme(plot.title = element_text(hjust = 0.5, size=14, face="bold.italic",colour = "grey38"))
ggsave("microbiome/output/ASV_Lactobacillus sp. NA.png", height = 8, width = 13)
```

### Snodgrassella
```{r Snodgrassella_ASV}
tax<-ps1
tax = rarefy_even_depth(tax, rngseed=1, sample.size=14240, replace=F)
ps2 <- subset_samples(tax, cycle =="cycle_one"|cycle =="cycle_two"|cycle =="cycle_three_before_stress")
ps2 <- subset_samples(ps2, treatment3 =="Control"|treatment3 =="Tetracycline")
ps2 <- subset_taxa(ps2, Genus =="Snodgrassella")
ps2.3 <- prune_taxa(taxa_sums(ps2) > 1000, ps2)
ps2m <- psmelt(ps2.3)
neworder <- c("cycle_one","cycle_two","cycle_three_before_stress")
change <- c("Cycle one","Cycle two","Cycle three")
ps_df_sum2 <- arrange(transform(ps2m, cycle=factor(cycle,levels=neworder,labels=change)),cycle)

p<-ggplot(data = ps_df_sum2, aes(x = treatment3, y = Abundance, colour = treatment3)) +
  geom_point(position = position_jitter())+geom_boxplot(show.legend=FALSE, alpha = 0,colour="grey53",size=0.7)+scale_color_manual(values=c("#55596a", "#6ebe9f"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7,alpha=0.5)))
p<-p + facet_wrap(~ASV, scales="free_y",nrow=2)
p<-p + theme(strip.text = element_text(size=13,face="bold",color="grey40"))
taxa2<-p+theme(axis.title.y = element_text(size=13, face="bold"))+theme(axis.text.y = element_text(size=12, face="bold"))+theme(axis.text.x = element_blank())+theme(axis.title.x = element_blank())+theme(strip.background =element_rect(fill="white", color="white"))
taxa2<- taxa2 + theme(legend.title = element_blank(),legend.background = element_blank(),legend.key = element_blank(),legend.text = element_text(size=14, face="italic"))
taxa2+ggtitle("Snodgrassella")+theme(plot.title = element_text(hjust = 0.5, size=14, face="bold.italic",colour = "grey38"))
ggsave("microbiome/output/ASV_Snodgrassella.png", height = 6, width = 10.5)
```

