---
title: "R_exprimental_survival_plots_and_analysis"
output: html_document
---


#Survival plots and analysis

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(GenomicFeatures)
library(dplyr)
library(survival)
library(km.ci)
library(magrittr)
library(ggpubr)
library(survminer)
library(gtable)
library(rlang)
library(tidyverse)
library(reshape2)
library(ggplot2)
library(fmsb)
library(Rmisc)
library(emmeans)
plot_res = 300 # DPI of PNG graphs
```

### plot mean of three cages per treatment with standard error
### Cycle 1

```{r cycle1, fig.width=6, fig.height=3}
cycle1 <- read.table("experiments_survival/cycle1_survival_2019.txt", header = TRUE)
tgc_cycle1 <- summarySE(cycle1, measurevar="percent_survived", groupvars=c("time","treatment"))

p<-ggplot(data=tgc_cycle1, aes(x=time, y=percent_survived, group = treatment, colour = treatment)) +
  geom_errorbar(aes(ymin=percent_survived-se, ymax=percent_survived+se), width=.1)+
  #geom_errorbar(aes(ymin=percent_survived-sd, ymax=percent_survived+sd), width=.3, size=1.1)+
  geom_line(size=1.5) +
  geom_point(size=2, shape=19, fill="white")

p<-p+scale_color_manual(values=c("control"="#55596a", "Tetracycline"="#D45E79"))
p<-p+coord_cartesian(xlim = c(0, 10))+scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))
p<-p+coord_cartesian(ylim=c(0, 1.1))+scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1))
p2<-p+theme(axis.line=element_line(colour="black"), panel.border=element_blank(),panel.background=element_blank())
p3<-p2+ theme(axis.text.y=element_text(size=15, face="bold"))+
  theme(axis.title.y=element_text(size=17))+theme(axis.title.x=element_text(size=17))+theme(axis.text.x=element_text(size=15, face="bold"))+
  #theme(legend.position = "bottomleft")
  theme(plot.title = element_text(size=15))+ggtitle("Cycle 1 survival")
p4<- p3+theme(legend.text = element_text(size=11))
p5 <- p4 + theme(legend.title = element_blank(),
                 legend.justification=c(1,0), 
                 legend.position=c(0.25, 0.05),  
                 legend.background = element_blank(),
                 legend.key = element_blank()) 
p_cycle1 <- p5 + theme(plot.title = element_text(hjust = 0.5)) +labs(x = "Time (days)")+labs(y = "Percent survival")
ggsave("experiments_survival/new_cycle1_survival_2019.png", height = 3, width = 6)
``` 

### plot mean of three cages per treatment with standard error
### Cycle 2

```{r cycle2, fig.width=6, fig.height=3, echo=FALSE, warning=FALSE}
cycle2 <- read.table("experiments_survival/cycle2_survival_2019.txt", header = TRUE)
tgc_cycle1 <- summarySE(cycle2, measurevar="percent_survived", groupvars=c("time","treatment"))

p<-ggplot(data=tgc_cycle1, aes(x=time, y=percent_survived, group = treatment, colour = treatment)) +geom_errorbar(aes(ymin=percent_survived-se, ymax=percent_survived+se), width=.1)+
  #geom_errorbar(aes(ymin=percent_survived-sd, ymax=percent_survived+sd), width=.3, size=1.1)+
  geom_line(size=1.5) +
  geom_point(size=2, shape=19, fill="white")
p<-p+scale_color_manual(values=c("control"="#55596a", "Tetracycline"="#D45E79"))
p<-p+coord_cartesian(xlim = c(0, 6))+scale_x_continuous(breaks=c(1,2,3,4,5,6))
p<-p+coord_cartesian(ylim=c(0, 1.1))+scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1))
p2<-p+theme(axis.line=element_line(colour="black"), panel.border=element_blank(),panel.background=element_blank())
p3<-p2+ theme(axis.text.y=element_text(size=15, face="bold"))+
  theme(axis.title.y=element_text(size=17))+theme(axis.title.x=element_text(size=17))+theme(axis.text.x=element_text(size=15, face="bold"))+
  #theme(legend.position = "bottomleft")
  theme(plot.title = element_text(size=15))+ggtitle("Cycle 2 survival")
p4<- p3+theme(legend.text = element_text(size=11))
p5 <- p4 + theme(legend.title = element_blank(),
                 legend.justification=c(1,0), 
                 legend.position=c(0.25, 0.05),  
                 legend.background = element_blank(),
                 legend.key = element_blank()) 
p_cycle2 <- p5 + theme(plot.title = element_text(hjust = 0.5)) +labs(x = "Time (days)")+labs(y = "Percent survival")
ggsave("experiments_survival/cycle2_survival_2019.png", height = 3, width = 6)
```
```{r cycle3, fig.width=6, fig.height=3, echo=FALSE, warning=FALSE}
cycle3 <- read.table("experiments_survival/cycle3_survival_before_stress_2019.txt", header = TRUE)
tgc_cycle3 <- summarySE(cycle3, measurevar="percent_survived", groupvars=c("time","treatment2"))

p<-ggplot(data=tgc_cycle3, aes(x=time, y=percent_survived, group = treatment2, colour = treatment2)) +geom_errorbar(aes(ymin=percent_survived-se, ymax=percent_survived+se), width=.1)+
  #geom_errorbar(aes(ymin=percent_survived-sd, ymax=percent_survived+sd), width=.3, size=1.1)+
  geom_line(size=1.5) +
  geom_point(size=2, shape=19, fill="white")
p<-p+scale_color_manual(values=c("control"="#55596a", "Tetracycline"="#D45E79"))
p<-p+coord_cartesian(xlim = c(0, 6))+scale_x_continuous(breaks=c(1,2,3,4,5,6))
p<-p+coord_cartesian(ylim=c(0, 1.1))+scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1))
p2<-p+theme(axis.line=element_line(colour="black"), panel.border=element_blank(),panel.background=element_blank())
p3<-p2+ theme(axis.text.y=element_text(size=15, face="bold"))+
  theme(axis.title.y=element_text(size=17))+theme(axis.title.x=element_text(size=17))+theme(axis.text.x=element_text(size=15, face="bold"))+
  #theme(legend.position = "bottomleft")
  theme(plot.title = element_text(size=15))+ggtitle("Cycle 3 survival")
p4<- p3+theme(legend.text = element_text(size=11))
p5 <- p4 + theme(legend.title = element_blank(),
                 legend.justification=c(1,0), 
                 legend.position=c(0.25, 0.05),  
                 legend.background = element_blank(),
                 legend.key = element_blank()) 
p_cycle3 <- p5 + theme(plot.title = element_text(hjust = 0.5)) +labs(x = "Time (days)")+labs(y = "Percent survival")
ggsave("experiments_survival/cycle3_survival_2019.png", height = 3, width = 6)
```

```{r arrange_survival_plots}

leg <- get_legend(p_cycle3)
legend2<-as_ggplot(leg)
ggsave("experiments_survival/legend2.png", height = 1, width = 5)

p_cycle3.1 <- p_cycle3 + theme(legend.position = "none",plot.title = element_blank())

p_cycle2.1 <- p_cycle2 + theme(legend.position = "none",axis.title.x = element_blank(), axis.title.y = element_blank(),plot.title = element_blank())

p_cycle1.1 <- p_cycle1 + theme(legend.position = "none",axis.title.x = element_blank(),plot.title = element_blank())

figure_supp_1 <- "experiments_survival/arrange_survival_plots_2019.png"
png(figure_supp_1, 13 * plot_res, 7 * plot_res, res = plot_res)
ggarrange(p_cycle1.1, p_cycle2.1, p_cycle3.1, ncol=2, nrow=2,labels = c("Cycle one", "Cycle two", "Cycle three"),font.label = list(size = 18, color = "dimgray"),vjust = c(1,1,1),hjust = c(-0.75,-0.5,-0.6))
invisible(dev.off())
knitr::include_graphics(figure_supp_1, dpi = plot_res)
```

```{r fisher_test}
tetra <- read.table("experiments_survival/tetra cycle 3 day 5 to 6 survival_2019.txt", header = TRUE)
fisher.test(tetra, alternative = "two.sided")

```



# brms survival functional tests

```{r setup2}
library(tidyverse)
library(brms)
library(tidybayes)
library(cowplot)
library(multcomp)
library(emmeans)
library(systemPipeR)
library(GOstats)
library(biomaRt)
library(GenomicFeatures)

dat <- read_tsv("experiments_survival/functional_test_Tetracycline_cycle3_2019.txt") %>% mutate(cage2 = paste(group, cage), group = factor(group, levels = c("control", "coevolved")))

controlDatFiltered <- read_tsv("experiments_survival/functional_test_filtered_guts_Tetracycline_2021.txt") %>% mutate(cage2 = paste(group, cage),group = factor(group, levels = c("filtered_control", "filtered_tetracycline")))
```

```{r cache=T}
priorsCage <- c(prior(student_t(3,0,2.5), class = "b"))
tet <- brm(alive ~ 0 + cage2 , data = filter(dat, treatment == "Tetracycline"), family= bernoulli(), control = list(adapt_delta = .99, max_treedepth = 10), iter = 5000, cores = 4, prior = priorsCage)
plot(tet)
summary(tet)
(hypTet <- hypothesis(tet, "cage2coevolvedcage1 + cage2coevolvedcage2 + cage2coevolvedcage3 = cage2controlcage1 +cage2controlcage2 + cage2controlcage3"))
```

# Filtered microbiome control

```{r cache=T}
ContrFiltered <- brm(alive ~ 0 + cage2, data = controlDatFiltered, family= bernoulli(), control = list(adapt_delta = .99), iter = 5000, cores = 4, prior = priorsCage)
plot(ContrFiltered)
summary(ContrFiltered)
(hypContrFiltered <- hypothesis(ContrFiltered, "cage2filtered_tetracyclinecage1 + cage2filtered_tetracyclinecage2 + cage2filtered_tetracyclinecage3 + cage2filtered_tetracyclinecage4  = cage2filtered_controlcage1 + cage2filtered_controlcage2 + cage2filtered_controlcage3 + cage2filtered_controlcage4"))
```

# Plot results

```{r}
p6 <- tibble("exposed\n   microbiome" = hypTet$samples[,1], "filtered guts" = hypContrFiltered$samples[,1]) %>% gather(key = "contrast", value = "b") %>% group_by(contrast) %>%  ggplot(aes(x = contrast, y=b)) + stat_halfeye(aes(fill=contrast),point_interval = mode_hdi, .width = c(.66, .95),alpha=0.5)+ geom_hline(yintercept = 0, color = "red") +scale_fill_manual(values=c("#D45E79","cadetblue"),guide = guide_legend(override.aes = list(shape = c(rep(15, 2)),size=7)))+ theme_minimal()+theme(axis.text.y = element_text(size=17))+theme(axis.title.x = element_blank(), axis.title.y = element_text(vjust = 0.5, size=16), plot.margin = unit(c(.5,.5,.5,.5), "cm"), axis.text.x = element_text(angle=45, hjust=1, size=18)) + coord_trans(ylim=c(-35,20)) + ylab(expression("Bayes factor survival difference"))+theme(legend.title = element_blank())+theme(legend.text = element_text(size=16,face="bold"))

p6
ggsave("experiments_survival/tetracycline_figure_manuscript.png", height = 5.5, width = 7.5)

#+theme(legend.position = "none")
```


### Statistics on survival data of bees with pre-exposed microbiomes vs respective controls under high chemical stress for main experiment

```{r survival_data}
tetra1 <- read.table("experiments_survival/fisher_survival_2019.txt", header = TRUE)
fisher.test(tetra1, alternative = "two.sided")

filtered <- read.table("experiments_survival/fisher_survival_filtered_gut_2021.txt", header = TRUE)
fisher.test(filtered, alternative = "two.sided")
```
